A matrix-free approach for finite-strain hyperelastic problems using geometric multigrid

### 2019-08-06 doc/large_strain.tex
move dPdF into appendix

{\color{red}
For completeness we mention that the fully referential linearization of the here considered constitutive equation \eqref{eq:neo_hookean_energy} results in the following equivalent bilinear form
}
\begin{align}
  A_{ij} \equiv a(\gz N_i, \gz N_j) =&
  \int_{\mcl B_0}
  \grad \gz N_i : D_{F}P : \grad \gz N_j
  \d V
  \label{eq:algebraic_tangent_dPdF}
  \\
  D_{F}P :=&
  2 \lambda
  \gz F^{-T} \otimes \gz F^{-T}
  +
  \left[\mu - 2 \lambda \ln\left( J \right) \right]
  \gz F^{-T} \underline{\otimes} \gz F^{-1}
  +
  \mu \gz I \overline{\otimes} \gz I
\end{align}
where the upper and lower dyadic products of pairs of second-order tensors are respectively given by
\begin{align*}
    \gz A \overline{\otimes} \gz B &= A_{ik} B_{jl} \gz e_i \otimes \gz e_j \otimes \gz e_k \otimes \gz e_l \\
    \gz A \underline{\otimes} \gz B &= A_{il} B_{jk} \gz e_i \otimes \gz e_j \otimes \gz e_k \otimes \gz e_l  \, .
\end{align*}

@@

>>>>

@@

\appendix

{\color{red}
It is a common practice to write bilinear form in fully referential configuration. For the here considered constitutive equation \eqref{eq:neo_hookean_energy} this results in the following bilinear form
}
\begin{align}
  A_{ij} \equiv a(\gz N_i, \gz N_j) =&
  \int_{\mcl B_0}
  \grad \gz N_i : {\rm D}_{\gz F}{\gz P} : \grad \gz N_j
  \d V
  \label{eq:algebraic_tangent_dPdF}
  \\
  {\rm D}_{\gz F}{\gz P} :=&
  2 \lambda
  \gz F^{-T} \otimes \gz F^{-T}
  +
  \left[\mu - 2 \lambda \ln\left( J \right) \right]
  \gz F^{-T} \underline{\otimes} \gz F^{-1}
  +
  \mu \gz I \overline{\otimes} \gz I
\end{align}
where the upper and lower dyadic products of pairs of second-order tensors are respectively given by
\begin{align*}
    \gz A \overline{\otimes} \gz B &= A_{ik} B_{jl} \gz e_i \otimes \gz e_j \otimes \gz e_k \otimes \gz e_l \\
    \gz A \underline{\otimes} \gz B &= A_{il} B_{jk} \gz e_i \otimes \gz e_j \otimes \gz e_k \otimes \gz e_l  \, .
\end{align*}

{\color{red}
This formulation can also be straight forwardly adopted within the matrix-free framework of \texttt{deal.II} library.
To that end a variation of the Algorithm \ref{alg:mf_tensor4} can be used where the gradients are evaluated with respect to the undeformed configuration and only a fourth-order referential tangent tensor ${\rm D}_{\gz F}{\gz P}$ is cached per quadrature point.
}

{\color{red}
In the interest of brevity, we did not report any results for this approach,
however we do support it in our code.
Below give a brief summary of the measured performance.
Clearly in this case we need to cache more data as compared to the Algorithm \ref{alg:mf_tensor4}, this results in both higher memory requirement per DoF as well as lower arithmetic intensity of the vmult operator.
On both architectures we recorded a slightly larger wall time per DoF as compared to the Algorithm \ref{alg:mf_tensor4}, however still much faster than the matrix-based approach.
From the LIKWID measurements we observed that this algorithm also results in lower performance on Cascade Lake.
The advantage of this algorithm is that it does not require evaluation of gradients in deformed configuration, therefore the matrix-free setup time is slightly lower as compared to the other three algorithms that rely on \texttt{MappingQEulerian} class of the \texttt{deal.II} library.
Overall we consider this approach to be a valid alternative to the Algorithm \ref{alg:mf_tensor4}
that can be be applied to any constitutive model which operates with the fourth-order referential tangent stiffness tensor on the quadrature point level.
}
